---
title: "DataAssignment_7"
author: "Md. Saifur Rahman"
date: "3/7/2021"
output: word_document
---

```{r setup the working directory, include=FALSE}
setwd("C:/Users/mdsaifur.rahman.1/Desktop/Spring 2021/EFR 535/Code R")

```

```{r load package, include=FALSE}
library(ggplot2)
library(tidyverse)
```

#  **Task-1: Load the Tophat dataset**

***

```{r load, view, and attach Tophat data}
library(readxl)
Data <- read_excel("EFR 535 Assign 2 - Data Tophat - Jan 18 2018.xls", 
    na = "NA")
View(Data)
attach(Data)
```

#  **Task-2:  Create a vertical bar chart for Year. Create a horizontal bar chart for Program of Study**

***

```{r Create bar chart}
# Create a vertical bar chart for Year.
ggplot(Data, aes(x=Year))+geom_bar()
# Create a horizontal bar chart for Program of Study.
ggplot(Data, aes(x=Program_of_Study))+geom_bar()+coord_flip()

```

#  **Task-3:  Change color of all bars**

***

```{r Change color of bars in bar chart}
# All bars with same color
ggplot(Data, aes(x=Year))+geom_bar(fill = "blueviolet")
ggplot(Data, aes(x=Program_of_Study))+geom_bar(fill = "darkcyan")+coord_flip()
# each bar has different color 
ggplot(Data, aes(x=Year, fill=Year))+geom_bar()
```

#  **Task-4:  Change the outline color of the bars so each is different according to the x-axis variable, leaving the fill white. You can use default colors, or the color brewer or grey scales**

***

```{r change outline color}
# fill white and outline color
ggplot(Data, aes(x=Year))+geom_bar(fill="white", color = "red")
ggplot(Data, aes(x=Program_of_Study))+geom_bar(fill="white", color = "chocolate")+coord_flip()
# Use of color brewer 
ggplot(Data, aes(x=Year,fill = Year,color = "red"))+geom_bar()+ scale_fill_brewer(palette = "YlOrRd")
ggplot(Data, aes(x=Program_of_Study, fill = Program_of_Study, color="red"))+geom_bar()+coord_flip()+ scale_fill_brewer(palette = "Spectral")
# Use of grey scales
ggplot(Data, aes(x=Year,fill = Year, color = "red"))+geom_bar()+ scale_fill_grey(start=0.10, end = 0.90)+ theme_classic()
ggplot(Data, aes(x=Program_of_Study, fill = Program_of_Study, color = "red"))+geom_bar()+coord_flip()+ scale_fill_grey(start=0.10, end = 0.90)+theme_classic()
```

#  **Task-5:  Change the fill color of the bars so each is different according to x-axis variable**

***

```{r Change fill color of bars in bar chart}
ggplot(Data, aes(x=Year, fill=Year, color="red"))+geom_bar()
ggplot(Data, aes(x=Program_of_Study,fill=Program_of_Study, color = "chocolate"))+geom_bar()+coord_flip()
```

#  **Task-6:   Adjust bar width and theme**

***

```{r bar width and theme}
# adjust bar width
ggplot(Data, aes(x=Year, fill=Year, color="red"))+geom_bar(width = 0.7)
ggplot(Data, aes(x=Program_of_Study,fill=Program_of_Study, color = "chocolate"))+geom_bar(width = 0.6)+coord_flip()
# adjust theme
ggplot(Data, aes(x=Year, fill=Year, color="red"))+geom_bar(width = 0.7)+ theme_classic()
ggplot(Data, aes(x=Program_of_Study,fill=Program_of_Study, color = "chocolate"))+geom_bar(width = 0.6)+coord_flip()+ theme_classic()
```

#  **Task-7:    Create a horizontal barplot in which categorical x-axis = Program of Study and continuous y-axis = Tophat_Qs_Correct. Then create vertical barplot in which categorical x-axis = Year and continuous y-axis = Tophat_total_Qs_answered.**

***

```{r create horizontal and vertical bar plot with both x-axis and y-axis variable with unsummarized data}
# Unsummarized data (seems wrong information in y-axis!!)
# Create a horizontal barplot in which categorical x-axis = Program of Study and continuous y-axis = Tophat_Qs_Correct.
ggplot(Data, aes(x=Program_of_Study, y=Tophat_Qs_Correct)) + geom_bar(stat="identity")
# I prefer to draw vertical barplot for this setup
ggplot(Data, aes(x=Program_of_Study, y=Tophat_Qs_Correct)) + geom_bar(stat="identity") + coord_flip()
# Then create vertical barplot in which categorical x-axis = Year and continuous y-axis = Tophat_total_Qs_answered.
ggplot(Data, aes(x=Year, y=Tophat_total_Qs_answered)) + geom_bar(stat="identity") + coord_flip()
# I prefer to draw horizontal barplot for this setup
ggplot(Data, aes(x=Year, y=Tophat_total_Qs_answered)) + geom_bar(stat="identity") 
```

#  **Task-8:    Create a new object that contains presummarized data for the variables in the bar charts. Change the color of the bars.**

***

```{r presummarized data}
# categorical x-axis = Program of Study and continuous y-axis = Tophat_Qs_Correct
means_CorrectPOS <- aggregate(Tophat_Qs_Correct ~ Program_of_Study, Data, FUN = mean)
view(means_CorrectPOS)
# Draw bar plot and change the colors of bars
ggplot(data=means_CorrectPOS, aes(x=Program_of_Study, y=Tophat_Qs_Correct, fill=Program_of_Study)) + geom_bar(stat="identity") + coord_flip()+scale_fill_brewer(palette = "Spectral")

# categorical x-axis = Year and continuous y-axis = Tophat_total_Qs_answered.
means_AnsweredYear <- aggregate(Tophat_total_Qs_answered ~ Year, Data, FUN = mean)
view(means_AnsweredYear)
# Draw bar plot and change the colors of bars
ggplot(data=means_AnsweredYear, aes(x=Year, y=Tophat_total_Qs_answered, fill=Year)) + geom_bar(stat="identity", width = 0.6) 
```

#  **Task-9:   Reorder the bars based on height**

***

```{r Reorder bar based on height}
# categorical x-axis = Program of Study and continuous y-axis = Tophat_Qs_Correct
ggplot(data=means_CorrectPOS, aes(x=reorder(Program_of_Study,-Tophat_Qs_Correct), y=Tophat_Qs_Correct, fill=Program_of_Study)) + geom_bar(stat="identity") + coord_flip()+scale_fill_brewer(palette = "Spectral")
# categorical x-axis = Year and continuous y-axis = Tophat_total_Qs_answered.
ggplot(data=means_AnsweredYear, aes(x=reorder(Year,Tophat_total_Qs_answered), y=Tophat_total_Qs_answered, fill=Year)) + geom_bar(stat="identity", width = 0.6) 
```

#  **Task-10:   Choose one of the plots above and create one variant of the bar chart: grouped barplot, dotplot, or lolliplot**

***

```{r gouped barchart/dotplot/lolliplot}
# Chosen barchart
ggplot(data=means_CorrectPOS, aes(x=reorder(Program_of_Study,-Tophat_Qs_Correct), y=Tophat_Qs_Correct, fill=Program_of_Study)) + geom_bar(stat="identity") + coord_flip()+scale_fill_brewer(palette = "Spectral")
# dotplot
ggplot(data=means_CorrectPOS, aes(x=Program_of_Study, y=Tophat_Qs_Correct)) + geom_point(size=4, color="red")+ theme_classic()+coord_flip()
# lolliplot
ggplot(data=means_CorrectPOS, aes(x=Program_of_Study, y=Tophat_Qs_Correct)) + geom_point(size=5, color="chocolate")+ geom_segment(aes(x = Program_of_Study, xend=Program_of_Study,y =0,yend=Tophat_Qs_Correct)) + theme_classic()+coord_flip()

```

